### 2017-03-20 paper/mapSkeletons.tex
minor

\subsection{parMapReduce}

>>>>

\subsection{parMapReduce}
-- this does not completely adhere to Google's definition of Map Reduce as it
-- the mapping function does not allow for "reordering" of the output
-- The original Google version can be found at https://de.wikipedia.org/wiki/MapReduce

\begin{lstlisting}[frame=htrbl]
parMapReduceDirect :: (ArrowParallel arr [a] b conf,
	ArrowApply arr, ArrowChoice arr) =>
	conf -> ChunkSize -> arr a b -> arr (b, b) b -> b -> arr [a] b
parMapReduceDirect conf chunkSize mapfn foldfn neutral =
	arr (chunksOf chunkSize) >>>
	parMap conf (mapArr mapfn >>> foldlArr foldfn neutral) >>>
	foldlArr foldfn neutral
\end{lstlisting}

### 2017-03-20 paper/topologySkeletons.tex
minor

\section{Topology Skeletons}
Even though many algorithms can be expressed by parallel maps, some problems require more sophisticated skeletons. The Eden library leverages this problem and already comes with a \textit{ring} and a \textit{torus} implementation \cite{eden_cefp, eden_skel_topology}. These two seem like reasonable candidates to be ported to our arrow based parallel Haskell to prove that its feasibility.

@@

@@

>>>>

\section{Topology Skeletons}
Even though many algorithms can be expressed by parallel maps, some problems require more sophisticated skeletons. The Eden library leverages this problem and already comes with more predefined skeletons, among them the \code{pipe}, \code{ring} and a \code{torus} implementation \cite{eden_cefp, eden_skel_topology}. These seem like reasonable candidates to be ported to our arrow based parallel Haskell to prove that we can express such skeletons with Arrows as well.

\subsection{pipe}

\begin{lstlisting}[frame=htrbl]
pipe :: (ArrowLoop arr, ArrowApply arr,
	ArrowParallel arr (fut a) (fut a) conf, Future fut a) =>
	conf -> [arr a a] -> arr a a
pipe conf fs = unliftFut $ pipeFut conf fs

pipeFut :: (ArrowLoop arr, ArrowApply arr,
	ArrowParallel arr (fut a) (fut a) conf, Future fut a) =>
	conf -> [arr a a] -> arr (fut a) (fut a)
pipeFut conf fs =
	resolve (arr $ \(a, outs) -> lazy $ a : outs) (parEvalNFut conf fs) >>>
	arr last
    where
        resolve :: (ArrowApply arr, ArrowLoop arr) =>
					arr (a, b) c -> arr c b -> arr a b
        resolve transform f =
					loop $ (arr $ \(a, b) -> (b, (f, (transform, (a, b))))) >>>
					second (second app >>> app)
\end{lstlisting}

@@

\begin{lstlisting}[frame=htrbl]
ring :: (ArrowLoop arr, ArrowApply arr, Future fut r, 
ArrowParallel arr (i, fut r) (o, fut r) conf) =>
	conf ->
	arr (i, r) (o, r) ->
	arr [i] [o]
ring conf f = loop $ second (arr rightRotate >>> arr lazy) >>>
	(arr $ uncurry zip) >>> (parMap conf (toFut $ f)) >>> arr unzip

-- similar to Eden's toRD
toFut :: (Arrow arr, Future fut r) =>
	arr (i, r) (o, r) ->
	arr (i, fut r) (o, fut r)
toFut f = (second $ arr get) >>> f >>> (second $ arr put)

-- from Eden:
rightRotate    :: [a] -> [a]
rightRotate [] =  []
rightRotate xs =  last xs : init xs
\end{lstlisting}

@@

\begin{lstlisting}[frame=htrbl]
torus :: (ArrowLoop arr, ArrowChoice arr, ArrowApply arr,
    ArrowParallel arr (c, fut [a], fut [b]) (d, fut [a], fut [b]) conf,
    ArrowParallel arr [(c, fut [a], fut [b])] [(d, fut [a], fut [b])] conf,
    Future fut [a], Future fut [b]) =>
	conf ->
	arr (c, [a], [b]) (d, [a], [b])	->
	arr [[c]] [[d]]
torus conf f =
	loop $ second (arr (map rightRotate) *** arr rightRotate) >>>
	arr (\ (inss, (inssA, inssB)) ->
		zipWith3 lazyzip3 inss (lazy inssA) (lazy inssB)) >>>
	parEvalNM conf (repeat (repeat (ptorus f))) >>>
	arr (map unzip3) >>> arr unzip3 >>> threetotwo

lazyzip3 :: [a] -> [b] -> [c] -> [(a, b, c)]
lazyzip3 as bs cs = zip3 as (lazy bs) (lazy cs)

ptorus :: (Arrow arr, Future fut [a], Future fut [b]) =>
          arr (c, [a], [b]) (d, [a], [b]) ->
          arr (c, fut [a], fut [b]) (d, fut [a], fut [b])
ptorus f =
	arr (\ ~(c, fas, fbs) -> (c, get fas, get fbs)) >>>
	f >>>
	arr (\ ~(c, as, bs) -> (c, put as, put bs))

threetotwo :: (Arrow arr) => arr (a, b, c) (a, (b, c))
threetotwo = arr $ \ ~(a, b, c) -> (a, (b, c))

parEvalNM :: (ArrowChoice arr, ArrowApply arr, ArrowParallel arr a b conf,
	ArrowParallel arr [a] [b] conf) =>
	conf -> [[arr a b]] -> arr [[a]] [[b]]
parEvalNM conf = listApp . (map (parEvalN conf))
\end{lstlisting}

### 2017-03-20 paper/utilityFunctions.tex
minor

\section{Utility Functions}\label{utilfns}
Before we go into detail on parallel arrows, we introduce some utility combinators first, that will help us later: map and zipWith on arrows.
\\
The mapArr combinator lifts any arrow \code{arr a b} to an arrow \code{arr [a] [b]} \cite{programming_with_arrows},
\begin{lstlisting}[frame=htrbl]
mapArr :: ArrowChoice arr => arr a b -> arr [a] [b]
mapArr f =
	arr listcase >>>
	arr (const []) ||| (f *** mapArr f >>> arr (uncurry (:)))
	where
		listcase [] = Left ()
		listcase (x:xs) = Right (x,xs)
\end{lstlisting}
and zipWithArr lifts any arrow \code{arr (a, b) c} to an arrow \code{arr ([a], [b]) [c]}.
\begin{lstlisting}[frame=htrbl]
zipWithArr :: ArrowChoice arr => arr (a, b) c -> arr ([a], [b]) [c]
zipWithArr f = (arr $ \(as, bs) -> zipWith (,) as bs) >>> mapArr f
\end{lstlisting}
These two combinators make use of the \code{ArrowChoice} typeclass, which allows us to use the \code{|||} combinator. It takes two arrows \code{arr a c} and \code{arr b c} and combines them into a new arrow \code{arr (Either a b) c} which pipes all \code{Left a}'s to the first arrow and all \code{Right b}'s to the second arrow.
\begin{lstlisting}[frame=htrbl]
(|||) :: ArrowChoice arr a c -> arr b c -> arr (Either a b) c
\end{lstlisting}
With the zipWithArr combinator we can also write a combinator \code{listApp}, which lifts a list of arrows \code{[arr a b]} to an arrow \code{arr [a] [b]}.
\begin{lstlisting}[frame=htrbl]
listApp :: (ArrowChoice arr, ArrowApply arr) => [arr a b] -> arr [a] [b]
listApp fs = (arr $ \as -> (fs, as)) >>> zipWithArr app
\end{lstlisting}
This combinator also makes use of the \code{ArrowApply} typeclass which allows us to evaluate arrows with \code{app :: arr (arr a b, a) c}.

>>>>

\section{Utility Functions}\label{utilfns}
Before we go into detail on parallel arrows, we introduce some utility combinators first, that will help us later: map, foldl and zipWith on arrows.
\\
The \code{mapArr} combinator lifts any arrow \code{arr a b} to an arrow \code{arr [a] [b]} \cite{programming_with_arrows}:
\begin{lstlisting}[frame=htrbl]
mapArr :: ArrowChoice arr => arr a b -> arr [a] [b]
mapArr f =
	arr listcase >>>
	arr (const []) ||| (f *** mapArr f >>> arr (uncurry (:)))
	where
		listcase [] = Left ()
		listcase (x:xs) = Right (x,xs)
\end{lstlisting}
Similarly, we can also define \code{foldlArr} that lifts any arrow \code{arr (b, a) b} with a neutral element \code{b} to \code{arr [a] b}:
\begin{lstlisting}[frame=htrbl]
foldlArr :: (ArrowChoice arr, ArrowApply arr) => arr (b, a) b -> b -> arr [a] b
foldlArr f b =
	arr listcase >>>
	arr (const b) |||
		(first (arr (\a -> (b, a)) >>> f >>> arr (foldlArr f)) >>> app)
	where
		listcase [] = Left []
		listcase (x:xs) = Right (x,xs)
\end{lstlisting}
Finally, with the help of \code{mapArr}, we can define zipWithArr, which lifts any arrow \code{arr (a, b) c} to an arrow \code{arr ([a], [b]) [c]}.
\begin{lstlisting}[frame=htrbl]
zipWithArr :: ArrowChoice arr => arr (a, b) c -> arr ([a], [b]) [c]
zipWithArr f = (arr $ \(as, bs) -> zipWith (,) as bs) >>> mapArr f
\end{lstlisting}
These combinators make use of the \code{ArrowChoice} typeclass, which allows us to use the \code{|||} combinator. It takes two arrows \code{arr a c} and \code{arr b c} and combines them into a new arrow \code{arr (Either a b) c} which pipes all \code{Left a}'s to the first arrow and all \code{Right b}'s to the second arrow.
\begin{lstlisting}[frame=htrbl]
(|||) :: ArrowChoice arr a c -> arr b c -> arr (Either a b) c
\end{lstlisting}
With the zipWithArr combinator we can also write a combinator \code{listApp}, which lifts a list of arrows \code{[arr a b]} to an arrow \code{arr [a] [b]}.
\begin{lstlisting}[frame=htrbl]
listApp :: (ArrowChoice arr, ArrowApply arr) => [arr a b] -> arr [a] [b]
listApp fs = (arr $ \as -> (fs, as)) >>> zipWithArr app
\end{lstlisting}
This combinator also makes use of the \code{ArrowApply} typeclass which allows us to evaluate arrows with \code{app :: arr (arr a b, a) c}.
